\documentclass[11pt,a4paper,oneside]{book}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{color}
\usepackage{calc}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage[pdftex]{hyperref}
% \hypersetup{pdftex, colorlinks=true, linkcolor=black, filecolor=black, urlcolor=black, citecolor=black,
	% pdftitle=Project Report, pdfauthor=Debjit Biswas}
\usepackage{amsmath}
\usepackage{setspace}

% \newlength{\myparindent}
% \setlength{\myparindent}{\parindent}
% \setlength{\parindent}{0in}
% \setlength{\topmargin}{0in}

\pagestyle{fancy}
\renewcommand{\chaptermark}[1]{%
        \markboth{#1}{}}
\renewcommand{\sectionmark}[1]{%
        \markright{\thesection\ #1}}
\fancyhf{}
\cfoot{\thepage}
\lhead{\bfseries\rightmark}
\rhead{\bfseries\leftmark}
\renewcommand{\headrulewidth}{0.5pt}
\renewcommand{\footrulewidth}{0pt}
\addtolength{\headheight}{0.5pt}
\fancypagestyle{plain}{%
   \fancyhead{} % get rid of headers on plain pages
   \renewcommand{\headrulewidth}{0pt} % and the line
}

\newcommand{\mycaption}[1]{\caption{\footnotesize #1} }

% \setlength{\parskip}{1.5ex plus 0.5ex minus 0.2ex}

% Paragraph styling
\setlength{\parindent}{0in}
\setlength{\parskip}{1\baselineskip}

% Line spacing
\onehalfspacing

%%%%%%%%%%%%%%%%%%%%%%%%%%%%Begin Document%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\input{frontmatter/title.tex}

\frontmatter

\input{frontmatter/certificate.tex}
\input{frontmatter/declaration.tex}

\tableofcontents

% \setlength{\parindent}{\myparindent}

\mainmatter

\input{mainmatter/chapter-1.tex}
\input{mainmatter/chapter-2.tex}
\input{mainmatter/chapter-3.tex}
\input{mainmatter/chapter-4.tex}
\input{mainmatter/chapter-5.tex}

\appendix

\input{mainmatter/appendix-a.tex}
\input{mainmatter/appendix-b.tex}

\normalsize
% \setlength{\parindent}{\myparindent}
\begin{thebibliography}{CCN$^+$05}

\bibitem[BZ96]{bz96} Jon Bennet and Hui Zhang. WF2 Q: Worst-case Fair Weighted Fair Queueing. 
  In \emph{Proceedings of INFOCOM ’96}, pages 120–128, 1996.

\bibitem[CCN$^+$05]{ccn05} Bogdan Caprita, Wong Chun Chan, Jason Nieh, Clifford Stein, and Haoquiang Zheng.
  Group Ratio Round Robin: O(1) Proportional Share Scheduling for Uniprocessor and
  Multiprocessor Systems. In \emph{Proceedings of the 2005 USENIX Annual Technical Conference}, pages 337–352, 2005.

\bibitem[DKS89]{dks89} Alan Demers, Srinivasan Keshav, and Scott Shenker. Analysis and Simulation of a Fair
  Queueing Algorithm. In \emph{Proceedings of ACM SIGCOMM ’89}, pages 1–12, 1989.

\bibitem[IEEE04]{ie04} The IEEE and The Open Group. \emph{The Open Group Base Specifications Issue 6}.
IEEE Std 1003.1, 2004 Edition. Available at http://www.opengroup.org/onlinepubs/009695399/.

\bibitem[Kle76]{kl76} Leonard Kleinrock. \emph{Queueing Systems, Volume 2: Computer Applications}.
Wiley-Interscience, First Edition, 1976. pp.~156-188.

\bibitem[Kol07]{ko07} Con Kolivas. \emph{RSDL completely fair starvation free interactive cpu scheduler}, 2007.
Available at http://lwn.net/Articles/224654/

\bibitem[Mol07]{mo07} Ingo Molnar. \emph{Modular Scheduler Core and Completely Fair Scheduler}, 2007.
Available at http://lwn.net/Articles/230501/

\bibitem[NVZ01]{nvz01} Jason Nieh, Chris Vaill, and Hua Zhong. Virtual Time Round Robin: An O(1) Propor-
  tional Share Scheduler. In \emph{Proceedings of the 2001 USENIX Annual Technical Confer-
  ence}, pages 245–259, 2001.


\bibitem[Tan01]{ta01} Andrew S.~Tanenbaum. \emph{Modern Operating Systems}.
Prentice~Hall, Second~Edition, 2001.

\bibitem[Tan06]{ta06} Andrew S.~Tanenbaum. \emph{Operating System Design and Implementation}. 
Prentice~Hall, Third~Edition, 2006.

\end{thebibliography}

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%Document End%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
